% Analysis 

```{r global_options, include = FALSE}
knitr::opts_chunk$set(cache = FALSE,
                      comment = "##",
                      collapse = TRUE,
                      warning = FALSE,
                      message = FALSE)

```

\renewcommand{\figurename}{Fig. S}

# Libraries

To reproduce the examples of this material, the R packages the following packages are needed.

```{r warning=FALSE, message=FALSE}
# it is suggested to use the dev version of metan package
# remotes::install_github("TiagoOlivoto/metan")

library(EnvRtype)
library(rio)
library(factoextra)
library(FactoMineR)
library(ggrepel)
library(ggh4x)
library(superheat)
library(ggridges)
library(metan)
library(rnaturalearth)
library(lme4)
library(lmerTest)
library(broom.mixed)
library(tidyverse)

# a ggplot2 theme for the plots
my_theme <- 
  theme_bw() +
  theme(panel.spacing = unit(0, "cm"),
        panel.grid = element_blank(),
        legend.position = "bottom")
```


# Dataset
```{r}
df_traits <- 
  import("data/df_traits.csv") |>
  metan::as_factor(1:6)


# long data
df_traits_long <- 
  df_traits |> 
  pivot_longer(GMC:HSW)


# grain yield mean in each environment
df_gy <-
  df_traits |>
  mean_by(ME, YEAR, .vars = GY)

# genotypic variance in each mega-environment
df_var_gy <- 
  df_traits |>
  group_by(YEAR, ME) |> 
  do(lmer(GY ~(1|GEN), data = .) |>
       tidy(effects = "ran_pars") |> 
       filter(group == "GEN") |> 
       transmute(var = estimate^2))

```


# Location map
```{r  fig.cap="Location map"}
china <- 
  ne_states(country = c("china", "taiwan"),
            returnclass = "sf")

locs <-
  import("https://bit.ly/local_info") %>%
  distinct(Lat, .keep_all = TRUE)

china <- 
  mutate(china,
         Province = ifelse(name %in% locs$Province, name, NA))


ggplot(data = china) +
  geom_sf(aes(fill = Province), size = 0.2) +
  ggthemes::theme_map() +
  scale_fill_discrete(na.value = "gray97",
                      labels = c(unique(locs$Province), "others")) +
  geom_point(data = locs,
             aes(x = Lon, y = Lat, size = Altitude),
             color = "black",
             fill = "red",
             shape = 21) +
  geom_label_repel(data = locs,
                   aes(label = Location,
                       x = Lon,
                       y = Lat),
                   fill = "green",
                   color = "black",
                   segment.color = 'black',
                   force = 38,
                   size = 3) +
  theme(legend.position = c(0.95, 0.1))
ggsave("figs/fig1_map.png", width = 10, height = 7)
```




# Mega-environment delineation
## 20-year climate data
```{r eval=FALSE}
df_years <- import("data/me_delineation.csv")
ENV <- df_years$Code
LAT <- df_years$Lat
LON <- df_years$Lon
ALT <- df_years$Alt
START <- df_years$Sowing
END <- df_years$Harvesting

# see more at https://github.com/allogamous/EnvRtype
df_climate <-
  get_weather(env.id = ENV,
              lat = LAT,
              lon = LON,
              start.day = START,
              end.day = END,
              parallel = TRUE)

# GDD: Growing Degree Day (oC/day)
# FRUE: Effect of temperature on radiation use efficiency (from 0 to 1)
# T2M_RANGE: Daily Temperature Range (oC day)
# SPV: Slope of saturation vapour pressure curve (kPa.Celsius)
# VPD: Vapour pressure deficit (kPa)
# ETP: Potential Evapotranspiration (mm.day)
# PEPT: Deficit by Precipitation (mm.day)
# n: Actual duration of sunshine (hour)
# N: Daylight hours (hour)


# RTA: Extraterrestrial radiation (MJ/m^2/day)
# T2M: Temperature at 2 Meters
# T2M_MAX: Maximum Temperature at 2 Meters
# T2M_MIN: Minimum Temperature at 2 Meters
# PRECTOT: Precipitation
# WS2M: Wind Speed at 2 Meters
# RH2M: Relative Humidity at 2 Meters
# T2MDEW: Dew/Frost Point at 2 Meters
# ALLSKY_SFC_LW_DWN: Downward Thermal Infrared (Longwave) Radiative Flux
# ALLSKY_SFC_SW_DWN: All Sky Insolation Incident on a Horizontal Surface
# ALLSKY_TOA_SW_DWN: Top-of-atmosphere Insolation
# [1] "env"       "ETP"       "GDD"         "PETP"      "RH2M"      "SPV"
# [8] "T2M"       "T2M_MAX"   "T2M_MIN"   "T2M_RANGE" "T2MDEW"    "VPD"

# Compute other parameters
env_data <-
  df_climate %>%
  as.data.frame() %>%
  param_temperature(Tbase1 = 10, # choose the base temperature here
                    Tbase2 = 33, # choose the base temperature here
                    merge = TRUE) %>%
  param_atmospheric(merge = TRUE) %>%
  param_radiation(merge = TRUE) |> 
  separate(env, into = c("env", "year"), sep = "_") |> 
  rename(TMAX = T2M_MAX,
         TMIN = T2M_MIN,
         ASKLW = ALLSKY_SFC_LW_DWN,
         ASKSW = ALLSKY_SFC_SW_DWN,
         TRANGE = T2M_RANGE)

```



```{r}
env_data <- readRDS("data/env_data.Rdata")

info <- 
  env_data |> 
  dplyr::select(env, LON, LAT) |> 
  mean_by(env)

df_gy_loc <-
  df_traits |>
  mean_by(LOC, .vars = GY)

env_wider <-  
  env_data |> 
  select(env, year, daysFromStart, TMED:RTA) |> 
  pivot_wider(names_from = "year",
              values_from = TMED:RTA) |> 
  left_join(info) |> 
  relocate(LON, LAT, .after = env) |> 
  mutate(MM = 1,
         DD = 1,
         DOY = 1,
         YYYYMMDD = 1,
         YEAR = 1,
         .after = env) |> 
  left_join(df_gy_loc, by = c("env" = "LOC"))

saveRDS(env_wider, "data/env_wider.Rdata")
```



## Environmental kinships

```{r}
id_var <- names(env_wider)[10:ncol(env_wider)]

EC <- 
  W_matrix(env.data = env_wider,
           env.id = "env",
           var.id = id_var,
           by.interval = TRUE,
           time.window = c(0, 30, 60, 90, 120, 150),
           QC = TRUE,
           sd.tol = 3)

saveRDS(EC, "data/EC.Rdata")
```



## Heatmap
```{r fig.width=8, fig.height=8, fig.cap="Similarity"}
distances <-
  env_kernel(env.data = EC,
             gaussian = TRUE)
d <-
  superheat(distances[[2]],
            heat.pal = c("#b35806", "white", "#542788"),
            pretty.order.rows = TRUE,
            pretty.order.cols = TRUE,
            col.dendrogram = TRUE,
            legend.width = 4,
            left.label.size = 0.1,
            bottom.label.text.size = 5,
            bottom.label.size = 0.2,
            bottom.label.text.angle = 90,
            legend.text.size = 17,
            heat.lim = c(0, 1),
            padding = 0.5,
            legend.height=0.2)

# ggsave(filename = "figs/fig2_heat_env2.pdf",
#        plot = d$plot,
#        width = 11,
#        height = 10)


env_data <- 
  env_data |> 
  mutate(me = case_when(
    env %in% c("Yicheng") ~ "ME1",
    env %in% c("Suixi", "Jieshou", "Nanyang") ~ "ME2",
    env %in% c("Shenzhou", "Gaocheng", "Handan", "Dezhou") ~ "ME3",
    env %in% c("Laizhou", "Jinan") ~ "ME4"
  ),
  .after = env)


```




## Principal component analysis

```{r fig.cap="biplot for PCA"}
env_data_m <- 
  env_data |> 
  select(-daysFromStart) |> 
  mean_by(me, .vars = TMED:RTA) |> 
  column_to_rownames("me")

# compute the PCA with
pca_model_h <- PCA(env_data_m,
                   graph = FALSE)

bp1 <- 
  fviz_pca_biplot(pca_model_h,
                  repel = TRUE,
                  col.var = "steelblue",
                  title = NULL) +
  coord_equal()


```



# Climate variables (trials)
## Getting the data

```{r eval=FALSE}
df_env <- import("data/location_info.csv")
ENV <- df_env$Code
LAT <- df_env$Lat
LON <- df_env$Lon
ALT <- df_env$Alt
START <- df_env$Sowing
END <- df_env$Harvesting

# see more at https://github.com/allogamous/EnvRtype
df_climate <-
  get_weather(env.id = ENV,
              lat = LAT,
              lon = LON,
              start.day = START,
              end.day = END)
# GDD: Growing Degree Day (oC/day)
# FRUE: Effect of temperature on radiation use efficiency (from 0 to 1)
# T2M_RANGE: Daily Temperature Range (oC day)
# SPV: Slope of saturation vapour pressure curve (kPa.Celsius)
# VPD: Vapour pressure deficit (kPa)
# ETP: Potential Evapotranspiration (mm.day)
# PEPT: Deficit by Precipitation (mm.day)
# n: Actual duration of sunshine (hour)
# N: Daylight hours (hour)
# RTA: Extraterrestrial radiation (MJ/m^2/day)
# SRAD: Solar radiation (MJ/m^2/day)
# T2M: Temperature at 2 Meters
# T2M_MAX: Maximum Temperature at 2 Meters
# T2M_MIN: Minimum Temperature at 2 Meters
# PRECTOT: Precipitation
# WS2M: Wind Speed at 2 Meters
# RH2M: Relative Humidity at 2 Meters
# T2MDEW: Dew/Frost Point at 2 Meters
# ALLSKY_SFC_LW_DWN: Downward Thermal Infrared (Longwave) Radiative Flux
# ALLSKY_SFC_SW_DWN: All Sky Insolation Incident on a Horizontal Surface
# ALLSKY_TOA_SW_DWN: Top-of-atmosphere Insolation
# [1] "env"       "ETP"       "GDD"         "PETP"      "RH2M"      "SPV"
# [8] "T2M"       "T2M_MAX"   "T2M_MIN"   "T2M_RANGE" "T2MDEW"    "VPD"

# Compute other parameters
env_data <-
  df_climate %>%
  as.data.frame() %>%
  param_temperature(Tbase1 = 10, # choose the base temperature here
                    Tbase2 = 33, # choose the base temperature here
                    merge = TRUE) %>%
  param_atmospheric(merge = TRUE) %>%
  param_radiation(merge = TRUE) |> 
  rename(DTIRF = ALLSKY_SFC_LW_DWN,
         SIHS = ALLSKY_SFC_SW_DWN,
         TMIN = T2M_MIN,
         TMAX = T2M_MAX, 
         TMED = T2M,
         RH = RH2M,
         WS = WS2M,
         TRANGE = T2M_RANGE) |> 
  separate(env, into = c("ENV", "YEAR")) |> 
  relocate(daysFromStart, .before = TMED) |> 
  mutate(ME = case_when(
    ENV %in% c("Yicheng") ~ "ME1",
    ENV %in% c("Suixi", "Jieshou", "Nanyang") ~ "ME2",
    ENV %in% c("Shenzhou", "Gaocheng", "Handan", "Dezhou") ~ "ME3",
    ENV %in% c("Laizhou", "Jinan") ~ "ME4"
  ),
  .after = ENV)

export(env_data, "data/env_data_trial.xlsx")
```


## Tidy climate data
```{r}
env_data_trial <- import("data/env_data_trial.xlsx")

```


## Principal component analysis

```{r fig.cap="biplot for PCA"}
names_var <- names(env_data_trial)[11:ncol(env_data_trial)]
pca1 <- 
  env_data_trial |> 
  mean_by(ME, YEAR, .vars = names_var) |> 
  metan::as_factor(1:2) |> 
  left_join(df_gy) |> 
  left_join(df_var_gy) |> 
  concatenate(ME, YEAR, .after = YEAR) |> 
  remove_cols(ME) |> 
  column_to_rownames("new_var")

pca2 <- 
  env_data_trial |> 
  mean_by(ME, .vars = names_var) |> 
  left_join(df_gy |> mean_by(ME)) |> 
  left_join(df_var_gy|> mean_by(ME)) |> 
  column_to_rownames("ME")


# compute the PCA with

pca_model <- PCA(pca1,
                 quali.sup = 1,
                 graph = FALSE)
pca_model2 <- PCA(pca2,
                  graph = FALSE)

bip1 <- 
  fviz_pca_biplot(pca_model,
                  repel = TRUE,
                  habillage = 1,
                  col.var = "steelblue",
                  title = NULL) +
  coord_equal()
bip2 <- 
  fviz_pca_biplot(pca_model2,
                  repel = TRUE,
                  col.var = "steelblue",
                  title = NULL) +
  coord_equal()


score_histori <- 
  pca_model_h[["ind"]][["coord"]] |> 
  as.data.frame() |>  
  rownames_to_column("me")

score_2a <- 
  pca_model2[["ind"]][["coord"]] |> 
  as.data.frame() |>  
  rownames_to_column("me")


arrange_ggplot(bp1, bip1, bip2, 
               tag_levels = "A")
ggsave("figs/fig3_pca.png", width = 20, height = 6.6)


```


## Environmental tipology (vapor pressure deficit)
```{r}
names.window <- c('1-intial growing','2-leaf expansion I','3-leaf expansion II','4-flowering','5-grain filling', "")


out <- 
  env_data_trial |> 
  concatenate(ME, YEAR, .after = YEAR, new_var = env) |> 
  env_typing(env.id = c("YEAR, ME"),
             var.id = names_var,
             by.interval = TRUE,
             time.window = c(0, 15, 35, 65, 90, 120),
             names.window = names.window,
             quantiles = c(.01, .25, .50, .75, .975, .99)) |> 
  separate(env, into = c("ME", "YEAR")) |> 
  separate(env.variable,
           into = c("var", "freq"),
           sep = "_",
           extra = "drop")

```


```{r fig.cap="Quantiles for vapor pressure deficit (A), deficit by precipitation (B), minimum air temperature (C), and maximum air temperature (D) observed in the studied mega-environments across distinct crop stages and cultivation years.", fig.width=13, fig.height=8}
# plot the distribution of envirotypes for dbp
variable <- "VPD"
p1 <-
  out |> 
  subset(var == variable) |> # change the variable here
  # mutate()
  ggplot() + 
  geom_bar(aes(x=Freq, y=ME,fill=freq), 
           position = "fill",
           stat = "identity",
           width = 1,
           color = "white",
           size=.2)+
  facet_grid(YEAR~interval, scales = "free", space = "free")+ 
  scale_y_discrete(expand = c(0,0))+
  scale_x_continuous(expand = c(0,0))+
  labs(x = 'Relative frequency',
       y = "Mega-Environment",
       fill='Envirotype')+
  theme(axis.title = element_text(size=12),
        legend.text = element_text(size=9),
        strip.text = element_text(size=12),
        legend.title = element_text(size=12),
        strip.background = element_rect(fill="gray95",size=1)) +
  ggthemes::scale_fill_stata()



# plot the distribution of envirotypes for dbp
variable <- "PETP"
p2 <-
  out |> 
  subset(var == variable) |> # change the variable here
  mutate(freq = as_factor(freq),
         freq = fct_relevel(freq, "(-13.1,-9.95]", "(-9.95,-7.54]",
                            "(-7.54,-2.13]", "(-2.13,31.5]",
                            "(31.5,47.8]" )) |> 
  ggplot() + 
  geom_bar(aes(x=Freq, y=ME,fill=freq), 
           position = "fill",
           stat = "identity",
           width = 1,
           color = "white",
           size=.2) +
  facet_grid(YEAR~interval) + 
  scale_y_discrete(expand = c(0,0))+
  scale_x_continuous(expand = c(0,0))+
  labs(x = 'Relative frequency',
       y = "Mega-Environment",
       fill='Envirotype')+
  theme(axis.title = element_text(size=12),
        legend.text = element_text(size=9),
        strip.text = element_text(size=12),
        legend.title = element_text(size=12),
        strip.background = element_rect(fill="gray95",size=1)) +
  ggthemes::scale_fill_stata()



# plot the distribution of envirotypes for dbp
variable <- "TMIN"
p3 <-
  out |> 
  subset(var == variable) |> # change the variable here
  # mutate()
  ggplot() + 
  geom_bar(aes(x=Freq, y=ME,fill=freq), 
           position = "fill",
           stat = "identity",
           width = 1,
           color = "white",
           size=.2) +
  facet_grid(YEAR~interval) + 
  scale_y_discrete(expand = c(0,0))+
  scale_x_continuous(expand = c(0,0))+
  labs(x = 'Relative frequency',
       y = "Mega-Environment",
       fill='Envirotype')+
  theme(axis.title = element_text(size=12),
        legend.text = element_text(size=9),
        strip.text = element_text(size=12),
        legend.title = element_text(size=12),
        strip.background = element_rect(fill="gray95",size=1)) +
  ggthemes::scale_fill_stata()




# plot the distribution of envirotypes for dbp
variable <- "TMAX"
p4 <-
  out |> 
  subset(var == variable) |> # change the variable here
  # mutate()
  ggplot() + 
  geom_bar(aes(x=Freq, y=ME,fill=freq), 
           position = "fill",
           stat = "identity",
           width = 1,
           color = "white",
           size=.2) +
  facet_grid(YEAR~interval) + 
  scale_y_discrete(expand = c(0,0))+
  scale_x_continuous(expand = c(0,0))+
  labs(x = 'Relative frequency',
       y = "Mega-Environment",
       fill='Envirotype')+
  theme(axis.title = element_text(size=12),
        legend.text = element_text(size=9),
        strip.text = element_text(size=12),
        legend.title = element_text(size=12),
        strip.background = element_rect(fill="gray95",size=1)) +
  ggthemes::scale_fill_stata()




arrange_ggplot(p1, p2,
               ncol = 1,
               tag_levels = "A")
ggsave("figs/fig4_typology_vpd.png", width = 10, height = 5)

```





# Helper functions
```{r}

# get variance components
get_vcomp <- function(model){
  model |> 
    tidy(effects = "ran_pars") |> 
    mutate(variance = estimate ^ 2) |> 
    dplyr::select(group, variance)
}

# get lrt
get_lrt <- function(model){
  model |> 
    ranova() |> 
    as.data.frame() |> 
    mutate(p_val = `Pr(>Chisq)`) |> 
    rownames_to_column("term") |> 
    dplyr::select(term, AIC, LRT, p_val) |> 
    mutate(term = str_sub(term, start = 6, end = -2))
}

```


# Models

```{r}

# fitted models
mod <- 
  df_traits_long |> 
  group_by(name) |> 
  doo(~ lmer(value ~ (1|ME) +   (1|GEN) + (1|GEN:YEAR) + (1|GEN:ME) + (1|GEN:ME:YEAR) + (1|YEAR/ME/REP2),
             data = .))

```


### lrt
```{r}
lrt <- 
  mod |> 
  mutate(ranova = map(data, ~.x |> get_lrt())) |> 
  dplyr::select(-data) |> 
  unnest(ranova)

lrt_wide <- 
  lrt |> 
  dplyr::select(name, term, p_val) |> 
  remove_rows_na() |>  
  pivot_wider(names_from = name, values_from = p_val)

export(lrt_wide, "data/lrt.xlsx")
```


## variance component
```{r}
# vcomp
vcomp <- 
  mod |> 
  mutate(vcomp = map(data, ~.x |> get_vcomp())) |> 
  dplyr::select(-data) |> 
  unnest(vcomp) |> 
  mutate(group = fct_relevel(group, "YEAR", "ME", "ME:YEAR", "GEN", "GEN:YEAR", "GEN:ME", "GEN:ME:YEAR", "REP2:(ME:YEAR)", "Residual"))

vcomp_wide <- 
  vcomp |> 
  pivot_wider(names_from = name, values_from = variance) |>
  rowwise() |> 
  mutate(d =  summarise(across(2:10, sum)))

ggplot(vcomp, aes(name, variance, fill = group)) +
  geom_bar(stat = "identity",
           position = "fill",
           color = "white",
           width = 1) +
  my_theme +
  theme(legend.position = "right") + 
  scale_x_discrete(expand = expansion(0)) +
  scale_y_continuous(expand = expansion(0),
                     labels = scales::label_percent()) +
  ggthemes::scale_fill_stata() +
  labs(x = "Traits",
       y = "Percentage of the total Variance (%)",
       fill = "Source")
ggsave("figs/fig5_vcomp.png", width = 6, height = 5)

```



# Selecion within ME
## Stability and mean performance
```{r}
waas <- 
  df_traits_long |> 
  dplyr::select(-REP) |> 
  pivot_wider(names_from = name, values_from = value) |> 
  group_by(ME) |> 
  doo(~mps(.,
           env = YEAR,
           gen = GEN,
           rep = REP2, 
           resp = GMC:HSW,
           ideotype_mper = c("l, l, l, l, h, h, h, h, h"),
           wmper = 70,
           performance = "blueg",
           stability = "ecovalence"))

```

## Join the WAASBY for all traits
```{r}
# mean performance and stability
mper <- 
  waas %>% 
  mutate(mps = map(data, ~.x %>% .[["mps_ind"]])) |> 
  unnest(mps) |> 
  dplyr::select(-data) |> 
  nest(data = GEN:HSW)

## MGIDI applied to the WAASBY values (all)
mgidi <- 
  mper |> 
  mutate(mgidi = map(data,
                     ~.x |> 
                       column_to_rownames("GEN") |> 
                       mgidi(SI = 25,
                             weights = c(1,1,1,1,5,1,1,1,1))))

p1 <- plot(mgidi$mgidi[[1]], SI = 25, y.lab = "MGIDI")
p2 <- plot(mgidi$mgidi[[2]], SI = 25, y.lab = "MGIDI")
p3 <- plot(mgidi$mgidi[[3]], SI = 25, y.lab = "MGIDI")
p4 <- plot(mgidi$mgidi[[4]], SI = 25, y.lab = "MGIDI")

arrange_ggplot(p1, p2, p3, p4,
               guides = "collect",
               tag_levels = "A",
               ncol = 4)

ggsave("figs/fig6_radar.png", width = 12, height = 4)

# selected genotypes
selm1 <- sel_gen(mgidi$mgidi[[1]])
selm2 <- sel_gen(mgidi$mgidi[[2]])
selm3 <- sel_gen(mgidi$mgidi[[3]])
selm4 <- sel_gen(mgidi$mgidi[[4]])


venn_plot(selm1, selm2, selm3, selm4,
          names = c("ME1", "ME2", "ME3", "ME4"),
          show_elements = TRUE) +
  ggthemes::scale_fill_stata()

ggsave("figs/fig7_veen.png", width = 6, height = 5)
```




## selection differentials for mean performance
```{r}
blue_mean <- 
  df_traits_long |> 
  mean_by(name, ME, GEN, .vars = value)

ovmean <- 
  df_traits_long |> 
  mean_by(name, ME, .vars = value)

# selected
selected <- 
  blue_mean |> 
  mutate(selected = case_when(
    ME == "ME1" & GEN %in% selm1 ~ "yes",
    ME == "ME2" & GEN %in% selm2 ~ "yes",
    ME == "ME3" & GEN %in% selm3 ~ "yes",
    ME == "ME4" & GEN %in% selm4 ~ "yes",
    TRUE ~ "no"
  )) |> 
  mean_by(name, ME, selected) |> 
  filter(selected == "yes")



ds_mper <- 
  ovmean |> 
  rename(xo = value) |> 
  left_join(selected |> rename(xs = value)) |> 
  dplyr::select(-selected) |> 
  mutate(sd_perc =  (xs - xo) / xo * 100,
         goal = case_when(
           name %in% c("BTL", "EH", "GMC", "PH") ~ "Negative desired gains",
           TRUE ~ "Positive desired gains"
         ),
         negative = ifelse(sd_perc <= 0, "Negative", "Positive"))

ggplot(ds_mper, aes(name, sd_perc)) +
  geom_hline(yintercept = 0, size = 0.2) +
  geom_col(aes(fill = negative),
           col = "black",
           size = 0.2) +
  scale_y_continuous(expand = expansion(mult = 0.2)) +
  facet_nested(~ME + goal, scales = "free_x") +
  geom_text(aes(label = round(sd_perc, 2),
                hjust = ifelse(sd_perc > 0, -0.1, 1.1),
                angle = 90),
            size = 3) +
  labs(x = "Traits",
       y = "Selection differential (%)") +
  my_theme +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        panel.grid.minor = element_blank()) +
  scale_fill_manual(values = ggthemes::stata_pal()(4)[c(2, 3)])

ggsave("figs/fig8_sd_mper.png", width = 14, height = 6)
```





## selection differentials for stability
```{r}

stab <- 
  waas %>% 
  mutate(stab = map(data, 
                    ~.x %>% .[["stability"]] |>
                      rownames_to_column("GEN"))) |> 
  unnest(stab) |> 
  dplyr::select(-data) |> 
  pivot_longer(GMC:HSW)


stab_ovmean <- 
  stab |> 
  mean_by(name, ME, .vars = value)

# selected
selected_stab <- 
  stab |> 
  mutate(selected = case_when(
    ME == "ME1" & GEN %in% selm1 ~ "yes",
    ME == "ME2" & GEN %in% selm2 ~ "yes",
    ME == "ME3" & GEN %in% selm3 ~ "yes",
    ME == "ME4" & GEN %in% selm4 ~ "yes",
    TRUE ~ "no"
  )) |> 
  mean_by(name, ME, selected) |> 
  filter(selected == "yes")



ds_stab <- 
  stab_ovmean |> 
  rename(xo = value) |> 
  left_join(selected_stab |> rename(xs = value)) |> 
  dplyr::select(-selected) |> 
  mutate(sd_perc =  (xs - xo) / xo * 100,
         negative = ifelse(sd_perc <= 0, "Negative", "Positive"))

ggplot(ds_stab, aes(name, sd_perc)) +
  geom_hline(yintercept = 0, size = 0.2) +
  geom_col(aes(fill = negative),
           col = "black",
           size = 0.2) +
  scale_y_continuous(expand = expansion(mult = 0.2)) +
  facet_nested(~ME , scales = "free_x") +
  geom_text(aes(label = round(sd_perc, 2),
                hjust = ifelse(sd_perc > 0, -0.1, 1.1),
                angle = 90),
            size = 3) +
  labs(x = "Traits",
       y = "Selection differential (%)") +
  my_theme +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        panel.grid.minor = element_blank()) +
  scale_fill_manual(values = ggthemes::stata_pal()(4)[c(2, 3)])
ggsave("figs/fig9_sd_stab.png", width = 14, height = 6)

```


# Rank winners in each mega-environment
```{r}
mean_yme <- 
  df_traits |> 
  mean_by(YEAR, ME, GEN, .vars = GY) |> 
  concatenate(ME, YEAR, new_var = ENV)

winners <- 
  ge_winners(mean_yme, ME, GEN, GY, type = "ranks") |> 
  mutate(id = rep(1:26, 4)) |>
  pivot_wider(names_from = ENV, values_from = GY)
```

# GGE
```{r}
mod_gge <- gge(mean_yme, ME, GEN, GY)

plot(mod_gge, type = 6)

ggsave("figs/fig10_gge.png", width = 10, height = 5)
```



# Section info
```{r}
sessionInfo()
```

