% Analysis 

```{r global_options, include = FALSE}
knitr::opts_chunk$set(cache = FALSE,
                      comment = "##",
                      collapse = TRUE,
                      warning = FALSE,
                      message = FALSE)

```

\renewcommand{\figurename}{Fig. S}

# Libraries

To reproduce the examples of this material, the R packages the following packages are needed.

```{r warning=FALSE, message=FALSE}
# it is suggested to use the dev version of metan package
# remotes::install_github("TiagoOlivoto/metan")

library(EnvRtype)
library(rio)
library(factoextra)
library(FactoMineR)
library(ggrepel)
library(ggh4x)
library(superheat)
library(ggridges)
library(tidyverse)
library(metan)
library(rnaturalearth)

# a ggplot2 theme for the plots
my_theme <- 
  theme_bw() +
  theme(panel.spacing = unit(0, "cm"),
        panel.grid = element_blank(),
        legend.position = "bottom")
```


# Datasets
## Traits
```{r}
df_traits <- 
  import("https://bit.ly/df_traits") |>
  metan::as_factor(1:3)


# grain yield mean in each environment
df_gy <-
  df_traits |>
  means_by(ENV, .vars = GY)



```


## Climate variables
### Scripts to gather data

```{r eval=FALSE}
df_env <- import("https://bit.ly/local_info")
ENV <- df_env$Code
LAT <- df_env$Lat
LON <- df_env$Lon
ALT <- df_env$Alt
START <- df_env$Sowing
END <- df_env$Harvesting

# see more at https://github.com/allogamous/EnvRtype
df_climate <-
  get_weather(env.id = ENV,
              lat = LAT,
              lon = LON,
              start.day = START,
              end.day = END)
# GDD: Growing Degree Day (oC/day)
# FRUE: Effect of temperature on radiation use efficiency (from 0 to 1)
# T2M_RANGE: Daily Temperature Range (oC day)
# SPV: Slope of saturation vapour pressure curve (kPa.Celsius)
# VPD: Vapour pressure deficit (kPa)
# ETP: Potential Evapotranspiration (mm.day)
# PEPT: Deficit by Precipitation (mm.day)
# n: Actual duration of sunshine (hour)
# N: Daylight hours (hour)
# RTA: Extraterrestrial radiation (MJ/m^2/day)
# SRAD: Solar radiation (MJ/m^2/day)
# T2M: Temperature at 2 Meters
# T2M_MAX: Maximum Temperature at 2 Meters
# T2M_MIN: Minimum Temperature at 2 Meters
# PRECTOT: Precipitation
# WS2M: Wind Speed at 2 Meters
# RH2M: Relative Humidity at 2 Meters
# T2MDEW: Dew/Frost Point at 2 Meters
# ALLSKY_SFC_LW_DWN: Downward Thermal Infrared (Longwave) Radiative Flux
# ALLSKY_SFC_SW_DWN: All Sky Insolation Incident on a Horizontal Surface
# ALLSKY_TOA_SW_DWN: Top-of-atmosphere Insolation
# [1] "env"       "ETP"       "GDD"         "PETP"      "RH2M"      "SPV"
# [8] "T2M"       "T2M_MAX"   "T2M_MIN"   "T2M_RANGE" "T2MDEW"    "VPD"

# Compute other parameters
env_data <-
  df_climate %>%
  as.data.frame() %>%
  param_temperature(Tbase1 = 10, # choose the base temperature here
                    Tbase2 = 33, # choose the base temperature here
                    merge = TRUE) %>%
  param_atmospheric(merge = TRUE) %>%
  param_radiation(merge = TRUE)
```

### Tidy climate data
```{r}
env_data_me <- import("https://bit.ly/df_climate")
env_data <- env_data_me |> remove_cols(me)
str(env_data)
```


# Scripts
## Location map
```{r  fig.cap="Location map"}
china <- 
  ne_states(country = c("china", "taiwan"),
            returnclass = "sf")

locs <-
  import("https://bit.ly/local_info") %>%
  distinct(Lat, .keep_all = TRUE)

china <- 
  mutate(china,
         Province = ifelse(name %in% locs$Province, name, NA))


ggplot(data = china) +
  geom_sf(aes(fill = Province), size = 0.2) +
  ggthemes::theme_map() +
  scale_fill_discrete(na.value = "gray97",
                      labels = c(unique(locs$Province), "others")) +
  geom_point(data = locs,
             aes(x = Lon, y = Lat, size = Altitude),
             color = "black",
             fill = "red",
             shape = 21) +
  geom_label_repel(data = locs,
                   aes(label = Location,
                       x = Lon,
                       y = Lat),
                   fill = "green",
                   color = "black",
                   segment.color = 'black',
                   force = 38,
                   size = 3) +
  theme(legend.position = c(0.95, 0.1))
ggsave("figs/fig1_map.png", width = 10, height = 7)
```


## Environmental kinships


```{r fig.width=8, fig.height=8, fig.cap="Similarity"}
id_var <- names(env_data)[10:20]

EC <- W_matrix(env.data = env_data,
               var.id = id_var,
               by.interval = TRUE,
               statistic = 'quantile',
               time.window = c(0, 15, 35, 65, 90, 120))
distances <-
  env_kernel(env.data = EC,
             gaussian = TRUE)
d <-
  superheat(distances[[2]],
            heat.pal = c("#b35806", "white", "#542788"),
            pretty.order.rows = TRUE,
            pretty.order.cols = TRUE,
            col.dendrogram = TRUE,
            legend.width = 4,
            left.label.size = 0.1,
            bottom.label.text.size = 5,
            bottom.label.size = 0.2,
            bottom.label.text.angle = 90,
            legend.text.size = 17,
            heat.lim = c(0, 1),
            padding = 0.5,
            legend.height=0.2)

ggsave(filename = "figs/fig2_heat_env.png",
       plot = d$plot,
       width = 10,
       height = 10)

```


```{r}

# long format for climate data
env_data_d <-
  env_data_me |>
  remove_cols(env, prec, LON:YYYYMMDD, daysFromStart) |>
  pivot_longer(-me)

# long format for grain yield
df_gy_dist <- 
  df_traits |> 
  select(me, GY) |> 
  mutate(name = "GY", .after = me) |> 
  rename(value = GY)

```


## Correlation analysis
```{r fig.cap="correlation analysis"}
env_data |> 
  corr_coef(tmean:sihs) |> 
  network_plot(min_cor = 0.4)

ggsave("figs/fig3_network.png", width = 8, height = 8)

```


## Principal component analysis

```{r fig.cap="biplot for PCA"}

pca <- 
  import("https://bit.ly/df_pca") |> 
  column_to_rownames("env")

# compute the PCA with

pca_model <- PCA(pca,
                 quali.sup = 14,
                 graph = FALSE)

fviz_pca_biplot(pca_model,
                repel = TRUE,
                habillage = 14,
                col.var = "gray40",
                title = NULL) +
  coord_equal()

ggsave("figs/fig4_pca.png", width = 7, height = 7)

```




## Environmental tipology (vapor pressure deficit)
```{r}
names.window <- c('1-intial growing','2-leaf expansion I','3-leaf expansion II','4-flowering','5-grain filling', "")

out <- 
  env_typing(env.data = env_data,
             env.id = "env",
             var.id = c("trange", "tmax", "tmin", "dbp", "etp", "vpd"),
             by.interval = TRUE,
             time.window = c(0, 15, 35, 65, 90, 120),
             names.window = names.window)

mes <- 
  distinct(df_traits, ENV, .keep_all = TRUE) |> 
  select(ENV, me)

out2 <- 
  separate(out,
           env.variable,
           into = c("var", "freq"),
           sep = "_",
           extra = "drop") |> 
  left_join(mes, by = c("env" = "ENV"))

```


```{r fig.cap="Quantiles for vapor pressure deficit observed in the studied environments (a) and mega-environments (b) across distinct crop stages.", fig.width=13, fig.height=7}
# plot the distribution of envirotypes for dbp
variable <- "vpd"
p1 <- 
  out2 |> 
  subset(var == variable) |> # change the variable here
  ggplot() + 
  geom_bar(aes(x=Freq, y=env,fill=freq), 
           position = "fill",
           stat = "identity",
           width = 1,
           color = "white",
           size=.2)+
  facet_grid(me~interval, scales = "free", space = "free")+ 
  scale_y_discrete(expand = c(0,0))+
  scale_x_continuous(expand = c(0,0))+
  xlab('Relative frequency of the vapor pressure deficit')+ 
  ylab("Environment")+
  labs(fill='Envirotype')+
  theme(axis.title = element_text(size=12),
        legend.text = element_text(size=9),
        strip.text = element_text(size=12),
        legend.title = element_text(size=12),
        strip.background = element_rect(fill="gray95",size=1),
        legend.position = 'bottom')

# by mega environment
p2 <- 
  out2 |> 
  subset(var == variable) |> # change the variable here
  sum_by(me, freq, interval) |> 
  ggplot() + 
  geom_bar(aes(x=Freq, y=me,fill=freq), 
           position = "fill",
           stat = "identity",
           width = 1,
           color = "white",
           size=.2)+
  facet_wrap(~interval, nrow = 1)+ 
  scale_y_discrete(expand = c(0,0))+
  scale_x_continuous(expand = c(0,0))+
  xlab('Relative frequency of the vapor pressure deficit')+ 
  ylab("Mega\nenvironment")+
  labs(fill='Envirotype')+
  theme(axis.title = element_text(size=12),
        legend.text = element_text(size=9),
        strip.text = element_text(size=12),
        legend.title = element_text(size=12),
        strip.background = element_rect(fill="gray95",size=1),
        legend.position = 'bottom') +
  scale_fill_discrete(direction = 1)

arrange_ggplot(p1, p2, 
               heights = c(0.85, 0.15),
               tag_levels = "A",
               guides = "collect")
ggsave("figs/fig5_typology_vpd.png", width = 12, height = 7)



```



## Selection for all environments
### Deviance analysis

```{r eval=FALSE}
mod_global <-
  waasb(df_traits,
        env = ENV,
        gen = GEN,
        rep = REP,
        resp = c(GMC, PH, EH, BTL, GY, EL, ER, GWE, HSW),
        mresp = c("l, l, l, l, h, h, h, h, h"),
        wresp = c(60, 60, 60, 60, 75, 60, 60, 60, 60),
        random = "all")
```

```{r echo=FALSE}
mod_global <- readRDS("data/mod_global.RData")
```

### MTSI for all the environments
```{r}
## mtsi global
SI <- 25
mtsi_global <- mtsi(mod_global, SI = SI)
sel_gen <- mtsi_global$sel_gen


```

## Selection within each mega-environment
### Mixed model

```{r eval=FALSE}
mod_me <-
  df_traits |> 
  waasb(env = ENV,
        gen = GEN,
        rep = REP,
        resp = c(GMC, PH, EH, BTL, GY, EL, ER, GWE, HSW),
        mresp = c("l, l, l, l, h, h, h, h, h"),
        wresp = c(60, 60, 60, 60, 75, 60, 60, 60, 60),
        random = "all",
        by = me)
```

```{r echo=FALSE}
mod_me <- readRDS("data/mod_me.RData")
```


### Multi-trait stability index

```{r}
## mtsi me1
mtsi_me1 <- mtsi(mod_me$data[[1]], SI = SI)
## mtsi me2
mtsi_me2 <- mtsi(mod_me$data[[2]], SI = SI)
## mtsi me3
mtsi_me3 <- mtsi(mod_me$data[[3]], SI = SI)

```

## Proportion of the phenotypic variance
```{r fig.width=10}
vcall <- plot(mod_global,
              type = "vcomp",
              percent.digits = 1,
              size.text.percent = 3,
              rotate = TRUE)
vc1 <- plot(mod_me$data[[1]],
            type = "vcomp",
            percent.digits = 1,
            size.text.percent = 3,
            rotate = TRUE)
vc2 <- plot(mod_me$data[[2]],
            type = "vcomp",
            percent.digits = 1,
            size.text.percent = 3,
            rotate = TRUE)
vc3 <- plot(mod_me$data[[3]],
            type = "vcomp",
            percent.digits = 1,
            size.text.percent = 3,
            rotate = TRUE)
arrange_ggplot(vcall, vc1, vc2, vc3,
               guides = "collect",
               tag_levels = "A")
ggsave("figs/fig6_vcomp.png", width = 15, height = 7)
```



## MTSI plot
```{r fig.width=10, fig.height=10, fig.cap="MTSI.", echo=FALSE}
m1 <- plot(mtsi_global, SI = 25)
m2 <- plot(mtsi_me1, SI = 25)
m3 <- plot(mtsi_me2, SI = 25)
m4 <- plot(mtsi_me3, SI = 25)

arrange_ggplot(m1, m2, m3, m4,
               tag_levels = "A",
               ncol = 2,
               guides = "collect")
ggsave("figs/fig7_radar.png", width = 7, height = 7)
```



## Selected genotypes in each ME
```{r fig.cap="selected genotypes"}
ALL <- mtsi_global$sel_gen
ME1 <- mtsi_me1$sel_gen
ME2 <- mtsi_me2$sel_gen
ME3 <- mtsi_me3$sel_gen
venn_plot(ALL, ME1, ME2, ME3, show_elements = TRUE)
ggsave("figs/fig8_venn.png", width = 5, height = 5)
```



## Selection gains for mean performance
```{r fig.cap="Selection gains for mean performance"}
sd_trait_global <-
  mtsi_global$sel_dif_trait %>%
  select_cols(VAR, sense, Xo, SDperc)

sd_me1 <- 
  mtsi_me1$sel_dif_trait |> 
  select_cols(VAR, SDperc) |> 
  rename(SD_me1 = SDperc)

sd_me2 <- 
  mtsi_me2$sel_dif_trait |> 
  select_cols(VAR, SDperc) |> 
  rename(SD_me2 = SDperc)

sd_me3 <- 
  mtsi_me3$sel_dif_trait |> 
  select_cols(VAR, SDperc) |> 
  rename(SD_me3 = SDperc)


sel_gain_joint <-
  sd_trait_global |>
  rename(SD_ALL = SDperc) |>
  left_join(sd_me1) |>
  left_join(sd_me2) |> 
  left_join(sd_me3)


df_gains <-
  sel_gain_joint |>
  remove_cols(Xo) |>
  pivot_longer(SD_ALL:SD_me3) |>
  replace_string(sense,
                 pattern = "increase",
                 replacement = "Positive desired gains") |>
  replace_string(sense,
                 pattern = "decrease",
                 replacement = "Negative desired gains") |>
  mutate(negative = ifelse(value <= 0, "Negative", "Positive")) |>
  replace_string(name, pattern = "SD_", replacement = "") |>
  tidy_strings(name, sep = "")



ggplot(df_gains, aes(VAR, value)) +
  geom_hline(yintercept = 0, size = 0.2) +
  geom_col(aes(fill = negative),
           col = "black",
           size = 0.2) +
  scale_y_continuous(expand = expansion(mult = 0.2)) +
  facet_nested(~name + sense, scales = "free_x") +
  geom_text(aes(label = round(value, 2),
                hjust = ifelse(value > 0, -0.1, 1.1),
                angle = 90),
            size = 3) +
  labs(x = "Selection strategy",
       y = "Selection differential (%)") +
  my_theme +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        panel.grid.minor = element_blank())

ggsave("figs/fig9_sd_mper.png", width = 14, height = 6)


# mean of differentials for BLUPs by each stragety
df_gains |>  means_by(name, negative)


```


## PC1 x PC2 and WAASB x Y
```{r fig.width=10, fig.height=4, fig.cap="PC1 x PC2 biplot"}
p1 <-
  plot_scores(mod_me$data[[1]],
              type = 2,
              var = "GY",
              highlight = ME1,
              size.tex.gen = 2.5,
              title = FALSE,
              col.alpha.gen = 0.6,
              col.gen = "gray",
              plot_theme = my_theme)

p2 <- 
  plot_scores(mod_me$data[[2]],
              type = 2,
              var = "GY",
              highlight = ME2,
              size.tex.gen = 2.5,
              title = FALSE,
              col.gen = "gray",
              col.alpha.gen = 0.6,
              plot_theme = my_theme)

p3 <- 
  plot_scores(mod_me$data[[3]],
              type = 2,
              var = "GY",
              highlight = ME3,
              size.tex.gen = 2.5,
              title = FALSE,
              col.gen = "gray",
              col.alpha.gen = 0.6,
              plot_theme = my_theme)

p4 <-
  plot_scores(mod_me$data[[1]],
              type = 3,
              var = "GY",
              highlight = ME1,
              x.lim = c(8, 13),
              x.lab = expression(Grain~yield~(t~ha^{-1})),
              size.tex.gen = 2.5,
              col.alpha.gen = 0.4,
              col.gen = "gray",
              size.bor.tick = 0.1,
              plot_theme = my_theme)

p5 <- 
  plot_scores(mod_me$data[[2]],
              type = 3,
              var = "GY",
              highlight = ME2,
              x.lab = expression(Grain~yield~(t~ha^{-1})),
              size.tex.gen = 2.5,
              x.lim = c(8, 13),
              col.alpha.gen = 0.4,
              col.gen = "gray",
              plot_theme = my_theme)

p6 <- 
  plot_scores(mod_me$data[[3]],
              type = 3,
              var = "GY",
              highlight = ME3,
              x.lab = expression(Grain~yield~(t~ha^{-1})),
              size.tex.gen = 2.5,
              x.lim = c(8, 13),
              col.alpha.gen = 0.4,
              col.gen = "gray",
              plot_theme = my_theme)

arrange_ggplot(p1, p2, p3,
               p4, p5, p6,
               guides = "collect",
               tag_levels = "A")
ggsave("figs/fig10_biplots.png", width = 14, height = 10)
```



## WAASB X Y biplot
```{r fig.width=10, fig.height=4, fig.cap="WAASB X Y biplot"}

```


# Section info
```{r}
sessionInfo()
```

